Review of Edge Detectors

### 2014-11-11 old_versions/edge-detection-review-2nd-revision-v1.tex
second round of third revision, improving references

The most usual approach to approximate the first derivative is to take the Taylor expansion of first order with a small $h$. Thus $f'(x)$ is computed as
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
\myn{\label{correction_a:first_derivative:discretization}} Digital images are commonly seen as a discretization of a continuos image, where a pixel is evenly sample on a grid. If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$. When no information about the spatial sampling is available, the size of the grid it is commonly assumed to be $\Delta x=\Delta y=1$.\\
Thus, in the discrete space the first derivative of the image is computed as:
\begin{equation}
\label{eq:roberts1}
	g_x = \frac{\partial f(x,y)}{\partial x} = f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j).\\
\end{equation}

\myn{\label{correction_a:first_derivative:diagonal edges}} Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$ by filtering the image $f(x,y)$ with the 1-D masks shown in Figure \ref{fig:1dmasks}. However, in order to be able to compute diagonal edges, 2-D masks are needed. \\
The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal \tcom{differences} \myn implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
usual to normalize the masks, in order to have gain equal to $1$. \\

@@

\myn{\label{correction_a:first_derivative:sobel} The left mask can be constructed (beyond normalization) as the convolution of a vertical derivative filter
$[-1\ 0\ 1]$ with an horizontally moving average filter $[1\ 1\ 1]$. The same reasoning could be applied to the right mask. Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter. Thus, they show a smoothing behaviour in the direction orthogonal to the gradient.

@@

Sobel masks can be seen (beyond normalization) as the convolution of a horizontal derivation mask 
$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph. \\

The computational difference between Prewitt and Sobel masks is not an issue. It is preferable to 
use Sobel operators, because they are better localized, and their filter are also less aliased, 
because of the weighted shape of $[1\ 2\ 1]$.

@@

The pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}.

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $tzc$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF
\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}
\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
\IF{Gaussian kernel}
	\STATE Define Laplacian approx. operator $laplacian$
	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
\ELSE
	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
\ENDIF
\STATE $max_L \leftarrow 0$
\FORALL{pixel $i$ in image $im_{LAPL}$}
	\IF{$im_{LAPL}[i]>max_L$}
		\STATE $max_L \leftarrow im_{LAPL}[i]$
	\ENDIF
\ENDFOR
\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
	\FORALL{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
		\IF{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \AND ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
			\STATE $im_{OUT}[i] \leftarrow 255$
		\ELSE
			\STATE $im_{OUT}[i] \leftarrow 0$
		\ENDIF
	\ENDFOR
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

The surrounding neighborhood of a point $(x,y)$ in the image $f$ is approximated using the following bi-cubic polynomial
\begin{equation}
	\label{eq:bicubic}
	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3. \\
\end{equation}

To solve this problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest 
neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.
In Haralick's work, this approximation is computed using least squares, and to speed up the calculations the solution is 
computed by convolving the image with some precomputed masks in Table \ref{table:b_i}, to find the coefficients $k_1\dots k_{10}$ at each point $(x,y)$. \\

Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
to find 10 coefficients. As mentioned before this leads to an overdetermied system which can be solved by least squares. 
By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
equations is obtained,

@@

\begin{equation*}
	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} \ \ \Rightarrow \ \ \mathbf{k} = \mathbf{B}\mathbf{f} .\\
\end{equation*}

$\mathbf{B}$ is a $10\times25$ matrix,

@@

\begin{equation}
	\label{eq:coefficients}
	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f}\star\mathbf{b_i},\\
\end{equation}

\begin{equation}
	\label{eq:b_i}
	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
									\vdots		& \vdots	& \ddots	& \vdots	\\
									b_{i,21}	& b_{i,22}	& \hdots	& b_{i,25}	\\
					\end{bmatrix}.\\
\end{equation}

\vspace{1ex}
Using equation \ref{eq:coefficients} and the masks given in Table \ref{table:b_i}, it is possible to compute 
the coefficients $k_1 \hdots k_{10}$ for all the points in the image. The elements of the mask $\mathbf{b_i}$ 
are the elements of the i-th row of $\mathbf{B} = (\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$.\\

\newcolumntype{C}{>{\centering\arraybackslash}m{1cm}<{}}
\begin{table}
\centering
\subfigure[$\mathbf{b_1}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
425 & 275 & 225 & 275 & 425 \\
\hline
275 & 125 &  75 & 125 & 275 \\ 
\hline
225 &  75 &  25 &  75 & 225 \\
\hline
275 & 125 &  75 & 125 & 275 \\
\hline
425 & 275 & 225 & 275 & 425 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_2}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-2260 & -620 & 0 & 620 & 2260 \\ 
\hline
-1660 & -320 & 0 & 320 & 1660 \\ 
\hline
-1460 & -220 & 0 & 220 & 1460 \\
\hline
-1660 & -320 & 0 & 320 & 1660 \\ 
\hline
-2260 & -620 & 0 & 620 & 2260 \\
\hline
\end{tabular}} \\
\subfigure[$\mathbf{b_3}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
2260  &  1660 &  1460 &  1660 &  2260 \\ 
\hline
620   &   320 &   220 &   320 &   620 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-620  &  -320 &  -220 &  -320 &  -620 \\ 
\hline
-2260 & -1660 & -1460 & -1660 & -2260 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_4}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
1130 & 620 & 450 & 620 & 1130 \\ 
\hline
830  & 320 & 150 & 320 &  830 \\ 
\hline
730  & 220 &  50 & 220 &  730 \\
\hline
830  & 320 & 150 & 320 &  830 \\ 
\hline
1130 & 620 & 450 & 620 & 1130 \\
\hline
\end{tabular}} \\
\subfigure[$\mathbf{b_5}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-400 & -200 & 0 &  200 &  400 \\ 
\hline
-200 & -100 & 0 &  100 &  200 \\ 
\hline
0    &    0 & 0 &    0 &    0 \\
\hline
200  &  100 & 0 & -100 & -200 \\ 
\hline
400  &  200 & 0 & -200 & -400 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_6}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
1130 & 830 & 730 & 830 & 1130 \\ 
\hline
620  & 320 & 220 & 320 &  620 \\ 
\hline
450  & 150 &  50 & 150 &  450 \\
\hline
620  & 320 & 220 & 320 &  620 \\ 
\hline
1130 & 830 & 730 & 830 & 1130 \\
\hline
\end{tabular}} \\
\subfigure[$\mathbf{b_7}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-8260 & -2180 & 0 & 2180 & 8260 \\ 
\hline
-6220 & -1160 & 0 & 1160 & 6220 \\ 
\hline
-5540 &  -820 & 0 &  820 & 5540 \\
\hline
-6220 & -1160 & 0 & 1160 & 6220 \\ 
\hline
-8260 & -2180 & 0 & 2180 & 8260 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_8}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
5640  &  3600 &  2920 &  3600 &  5640 \\ 
\hline
1800  &   780 &   440 &   780 &  1800 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-1800 &  -780 &  -440 &  -780 & -1800 \\ 
\hline
-5640 & -3600 & -2920 & -3600 & -5640 \\
\hline
\end{tabular}} \\
\subfigure[$\mathbf{b_9}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
-5640 & -1800 & 0 & 1800 & 5640 \\ 
\hline
-3600 &  -780 & 0 &  780 & 3600 \\ 
\hline
-2920 &  -440 & 0 &  440 & 2920 \\
\hline
-3600 &  -780 & 0 &  780 & 3600 \\ 
\hline
-5640 & -1800 & 0 & 1800 & 5640 \\
\hline
\end{tabular}}
\qquad\qquad
\subfigure[$\mathbf{b_{10}}$.] {
\begin{tabular}{|*{5}{C|}}
\hline
8260  &  6220 &  5540 &  6220 &  8260 \\ 
\hline
2180  &  1160 &   820 &  1160 &  2180 \\ 
\hline
0     &     0 &     0 &     0 &     0 \\
\hline
-2180 & -1160 &  -820 & -1160 & -2180 \\ 
\hline
-8260 & -6220 & -5540 & -6220 & -8260 \\
\hline
\end{tabular}} \\
\caption{Masks to compute the coefficients of the bicubic fit.}
\label{table:b_i}
\end{table}

@@

The neighborhood of each point of the image is approximated using the bi-cubic polynomial 
expression in equation \ref{eq:bicubic}. If just the first order terms of this 
polynomial are taken, the gradient angle $\theta$, defined with negative x-axis, can be approxmated as

@@

Now substituting the variables $x$ and $y$ in polar form as

@@

\begin{equation}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0 \ \ \Rightarrow \ \ \left| \frac{C_2}{3C_3} \right| < \rho_0,
\end{equation}

where $\rho_0$ is a fixed threshold, passed as an argument to the algorithm. Ideally, the polynomial function 
should be zero at $\rho=0$, but this condition is relaxed, allowing the the polynomial becomes zero in a neighborhood 
of $\rho=0$ ($\rho\leq\rho_0$). This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal 
value is between $0.4$ and $0.6$.\\

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF
\STATE Define the ten $5x5$ masks ($mask[1]\dots mask[10]$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

All algorithms implemented use some common mathematical operations that are independent of the algorithms themselves. 
These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
implemented with care. In this section those operations are reviewed.

@@

\section{Implementation}
\label{sec:appendix2}

@@

Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs significative faster that the LoG one. The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel). It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels. \\

Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas. It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic polinomial function. This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too. \\

>>>>

\myn{\label{correction_a:first_derivative:taylor}}The most usual approach to approximate the first derivative is to use a the first order Taylor expansion with a small value of $h$. Thus $f'(x)$ is computed as
\begin{equation}
	f'(x) \simeq \frac{f(x+h) - f(x)}{h}.\\
\end{equation}
\myn{\label{correction_a:first_derivative:discretization}} Digital images are commonly seen as a discretization of a continuos image, where a pixel is evenly sample on a grid. If we denote the grid sizes by $\Delta x$ and $\Delta y$, the continuous coordinates $x$  and $y$ can be written as $x=i\Delta x$ and $y=i\Delta y$. When no information about the spatial sampling is available, the size of the grid it is commonly assumed to be $\Delta x=\Delta y=1$.\\
Thus, in the discrete space the first derivative of the image is computed as:
\begin{equation}
\label{eq:roberts1}
	g_x = \frac{\partial f(x,y)}{\partial x} = f(i+1,j) - f(i,j)
\end{equation}
and
\begin{equation}
\label{eq:roberts2}
	g_y = \frac{\partial f(x,y)}{\partial y} = f(i,j+1) - f(i,j).\\
\end{equation}

\myn{\label{correction_a:first_derivative:diagonal edges}} Equations \ref{eq:roberts1} and \ref{eq:roberts2} can be implemented for all values of $x$ and $y$ by filtering the image $f(x,y)$ with the 1-D masks shown in Figure \ref{fig:1dmasks}. However, in order to be able to compute diagonal edges, 2-D masks are needed. \\
The \textit{Roberts operators} are one of the earliest 
attempts to use 2-D masks for this purpose. These operators are based on computing the diagonal \tcom{differences} \myn{\label{correction_a:typo}} implemented by filtering an image with the masks in Figure \ref{fig:roberts}. By convention 
in these figures, x-coordinates grow from left to right and y-coordinates grow top-down. It is also 
usual to normalize the masks, in order to have gain equal to $1$. \\

@@

\myn{\label{correction_a:first_derivative:sobel}} The left mask can be constructed (up to the normalization factor) as the convolution of two filters: a vertical derivative 
$[-1\ 0\ 1]$ and an horizontal moving average $[1\ 1\ 1]$. The same reasoning could be applied to the right mask. Hence, these operators can be interpreted as the consecutive application of an horizontal low pass filter with a vertical derivative filter. Thus, they show a smoothing behaviour in the direction orthogonal to the gradient.

@@

Sobel masks can be seen (up to the normalization factor) as the convolution of a horizontal derivation mask 
$[-1\ 0\ 1]$ with a vertical smoothing filter $[1\ 2\ 1]$ (closer to a Gaussian response than Prewitt), and vice versa. 
Hence, these operators have better smoothing properties, as mentioned in the preceding paragraph. \\

The computational cost of applying Prewitt and Sobel masks is exactly the same. Thus, it is preferable to 
use Sobel operators, because they are better localized, and their filter are also less aliased, because of the weighted shape of $[1\ 2\ 1]$.

@@

The pseudo-code of Marr-Hildreth's algorithm is shown in Algorithm \ref{algo:mh}

\begin{algorithm}[t!]
\caption{Marr-Hildreth edge detection algorithm.}
\label{algo:mh}
\begin{algorithmic}[1]
\REQUIRE input image, standard deviation $\sigma$, kernel size $n$ and zero-crossing threshold $th_{ZC}$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF

\IF{Gaussian kernel}
	\STATE Define Laplacian approx. operator $laplacian$
	\STATE $im_{LAPL} \leftarrow$ convolution($im$,$laplacian$)
\ELSE
	\STATE $kernel \leftarrow$ generate\_kernel($n$,$\sigma$) \COMMENT{Generated Gaussian or LoG kernel}\mynf{\label{correction_a:second_derivative:marr_algo}}.
	\STATE $im_{SMOOTHED} \leftarrow$ convolution($im$,$kernel$)
	\STATE $im_{LAPL} \leftarrow im_{SMOOTHED}$
\ENDIF
\STATE $max_L \leftarrow 0$
\FORALL{pixel $i$ in image $im_{LAPL}$}
	\IF{$im_{LAPL}[i]>max_L$}
		\STATE $max_L \leftarrow im_{LAPL}[i]$
	\ENDIF
\ENDFOR
\FORALL{pixel $i$ in image $im_{LAPL}$, except borders}
	\FORALL{pair $(p_1,p_2)$ of opposite neighbors of $p$ in $im_{LAPL}$}
		\IF{($\mbox{sign}(im_{LAPL}[p_1])\neq\mbox{sign}(im_{LAPL}[p_2])$) \AND ($|im_{LAPL}[p_1]-im_{LAPL}[p_2]|>th_{ZC}$)}
			\STATE $im_{OUT}[i] \leftarrow 255$
		\ELSE
			\STATE $im_{OUT}[i] \leftarrow 0$
		\ENDIF
	\ENDFOR
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

The surrounding neighborhood of a point $(x,y)$ in the image $f$ is approximated using the following bi-cubic polynomial
\begin{equation}
	\label{eq:bicubic}
	f(x,y) = k_1 + k_2x + k_3y + k_4x^2 + k_5xy + k_6y^2 + k_7x^3 + k_8x^2y + k_9xy^2 + k_{10}y^3. \\
\end{equation}
To solve this fitting problem, it is necessary to take more neighbors than coefficients to be adjusted. As there are 10 coefficients to compute, the smallest neighborhood of odd size that accomplishes this has size $5\times5$. Having $10$ coefficients and $25$ data 
points leads to an overdetermined system, which can be solved using least squares or any other fitting technique.\\

\myn{\label{second_derivative:haralick:statement}} In Haralick's work, this polynomial approximation is expressed in a local coordinate system centered on each pixel. In this way, the center of the neighboorhood is always $(0,0)$, and the coordinates $(x, y)$ are offsets from this center point (e.g. in a neighborhood of size 5x5, x and y take values between -2 and +2).\\
Then the solution is computed using least squares, and the local expression of the polynomial ensures that the values of the $k_i$ coefficientes are computed in the same way for every pixel in the image. Thus, they can be computed by convolving the image with a set of fixed and precomputed masks, which is a great speed-up of the algorithm. In the following we will state the approximation problem and show how to derive those masks from its solution.\\

Consider 25 points in a small neighborhood of a point $(x,y)$ in the image. This gives us an equal number of equations 
to find 10 coefficients. As mentioned before this leads to an \tcom{overdetermined} system which can be solved by least squares. 
By substituting the 25 data points into the polinomial equation (\ref{eq:bicubic}), the following system of 
equations is obtained,

@@

\begin{equation*}
	(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{f} = \mathbf{k} 
\end{equation*}

\myn{\label{second_derivative:haralick:normal}} Let us define a matrix $\mathbf{B} \in R^{10\times25}$ such that $\mathbf{B}=(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T$, thus

$$ \mathbf{k} = \mathbf{B}\mathbf{f} $$

@@

\begin{equation}
	\label{eq:coefficients}
	k_i = b_{i,1}f_1 + b_{i,2}f_2 + b_{i,3}f_3 + \hdots + b_{i,25}f_{25} \ \ \Rightarrow \ \ \mathbf{k_i} = \mathbf{f} . \mathbf{b_i},\\
\end{equation}

where $b_i$ is the vector containing the $i-th$ row of $B$, i.e.\myn{\label{correction_a:second_derivative:haralick_bi}}

\begin{equation}
	\label{eq:b_i_v}
	\mathbf{b_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,25}
					\end{bmatrix}.\\
\end{equation}

Let us rearrange $b_i$ as a 5x5 matrix and denote it by $\mathbf{mask_i}$ as follows
\begin{equation}
	\label{eq:b_i}
	\mathbf{mask_i} = \begin{bmatrix}	b_{i,1}		& b_{i,2}	& \hdots	& b_{i,5}	\\
									b_{i,6}		& b_{i,7}	& \hdots	& b_{i,10}	\\
									\vdots		& \vdots	& \ddots	& \vdots	\\
									b_{i,21}	 & b_{i,22}	& \hdots	& b_{i,25}	\\
					\end{bmatrix}.\\
\end{equation}

\vspace{1ex}
\myn{\label{second_derivative:haralick:mask_computation}} We will compute one value of $k_i$ for each pixel on the image. Thus we can store these coefficientes in a matrix $\bf{K_i}$ of the same size as the original image. Then, using the mask in shown in Equation \ref{eq:b_i} we can obtain $\bf{K_i}$ as the convolution of the original image with $\mathbf{mask_i}$. \\
\begin{equation}
	\label{eq:convolution_b}
	\mathbf{K_i} = f \star \mathbf{mask_i} \mathtt{for} i \in {1 \ldots 10}\\
\end{equation}

As we expressed the approximation in a local coordinate system, the coefficients $b_{i,j}$ will be the same for every pixel, thus, this convolution is performed against a fixed and precomputed mask $\mathbf{mask_i}$. The numeric values of these 5x5 masks used in our experiments are shown in Table \ref{table:b_i}.

@@

\myn{\label{second_derivative:haralick:analitical:statement}} After the neighborhood of each point of the image is approximated using the bi-cubic polynomial expression in equation \ref{eq:bicubic}, the idea of Haralick's algorithm is to compute the second derivative of the image in the direction of the gradient using that approximation.\\

Let $\theta$ be the gradient angle, defined with respect to the negative x-axis, and it is \tcom{approximated} using the first order terms of the 
polynomial \ref{eq:bicubic} obtaining the following expression:

@@

\myn{\label{second_derivative:haralick:analitical:derivative}} Let us consider the (half) line $r$ passing by $(x,y)$ with angle $\theta$. Each point $(u,v)$ on this line can be described in terms of  the angle $\theta$ and the distance $\rho$ to $(x,y)$ as follows

@@

\begin{equation}\label{eq:haralick:cond_b_orig}
	f''_{\theta}(\rho) = 2C_2 + 6C_3\rho = 0
\end{equation}

 Ideally, to hold this condition, the polynomial function 
should be zero at $\rho=0$, but this condition is relaxed, allowing the polynomial becomes zero in a neighborhood 
of $\rho=0$ ($\rho\leq\rho_0$). \\
With this relaxation, the condition \ref{eq:haralick:cond_b_orig} becomes

\begin{equation}
\left| \frac{C_2}{3C_3} \right| < \rho_0,
\end{equation}

where $\rho_0$ is a fixed threshold, which passed as an argument to the algorithm.
\myn{\label{second_derivative:haralick:analitical:parameter_value}}  This parameter must be greater than zero and and less than $\sqrt{2}$. Its optimal  value is between $0.4$ and $0.6$.\\

@@

\begin{algorithm}[t]
\caption{Haralick edge detection algorithm.}
\label{algo:haralick}
\begin{algorithmic}[1]
\REQUIRE input image (width $w$, height $h$), edge point condition threshold $\rho_0$.
\STATE $im_{ORIG} \leftarrow$ input image
\IF{$im_{ORIG}$ is grayscale}
	\STATE $im \leftarrow im_{ORIG}$
\ELSE 
	\STATE $im \leftarrow (6968im_{ORIG,R}+23434im_{ORIG,G}+2366im_{ORIG,B})/32768$ \COMMENT{Grayscale conversion, assuming RGB image.}
\ENDIF
\STATE Define the ten $5x5$ masks ($mask[1] \dots mask[10]$) used to determine coefficients $k_1$ to $k_{10}$ \COMMENT{See Table \ref{table:b_i}.} \mynf{\label{correction_a:second_derivative:haralick_algo}}
\FORALL{pixel $i$ in image $im$}
	\STATE $neighbors \leftarrow$ $5\times5$ neighborhood of pixel $i$ in image $im$
	\FOR{$j=1$ \TO $10$}
		\STATE $k_j \leftarrow$ convolution($neighbors$,$mask[j]$)
	\ENDFOR
	\STATE $C_2 \leftarrow \frac{k_2^2k_4 + k_2k_3k_5 + k_3^2k_6}{k_2^2 + k_3^2}$
	\STATE $C_3 \leftarrow \frac{k_2^3k_7 + k_2^2k_3k_8 + k_2k_3^2k_9 + k_3^3k_{10}}{(\sqrt{k_2^2 + k_3^2})^3}$
	\IF{{$|C_2/3C_3|\leq\rho_0$} \AND {$C_3<0$}}
		\STATE $im_{OUT}[i] \leftarrow 255$
	\ELSE
		\STATE $im_{OUT}[i] \leftarrow 0$
	\ENDIF
\ENDFOR
\RETURN output image $\leftarrow im_{OUT}$
\end{algorithmic}
\end{algorithm}

@@

\myn{\label{}} All implemented algorithms use some common mathematical operations that are independent of the algorithms themselves. 
These operations, although basic, have a great impact on the algorithm outcome, and thus they need to be 
implemented with care. In this section those operations are reviewed.

@@

@@

Comparing the two versions of the Marr-Hildreth algorithm, the version with Gaussian kernel runs \tcom{significantly} faster that the LoG one. The latter, while slower (since it needs a larger kernel to achieve similar results) is more accurate, because it makes no approximation to calculate the Laplacian (it is calculated analytically before creating the kernel). It is also possible to manage the size of the kernel, which represents a scale parameter of the algorithm, being able to obtain a highly detailed edge image using small kernels, or just more noticeable edges using larger kernels. \\

Haralick's algorithm, although it shares the Marr-Hildreth idea of finding zero crossings of the second derivative, has some quite different ideas. It is the only one of these algorithms which works with an approximation of the intensity of the image in the neighborhood of a pixel, using a bicubic \tcom{polynomial} function. This supposes a fairly large regularity in the image, which is not always true, and sometimes leads to detect edges where none of them exist, and get some thicker edges too. \\

### 2014-11-11 old_versions/edge-detection-review-3rd-revision-v1.tex
second round of third revision, improving references

Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts \cite{im_proc:segmentation:roberts:1963:boundary_extraction} proposes
to compute them as shown in Figure \ref{fig:roberts}. In order to use the same convolution code for all algorithms, our implementation uses $3\times3$ matrices, adding a row and a column of zeros to the matrices in Figure \ref{fig:roberts}, what produces the same result.

@@

%-------------------------------------------------------------------------------
\clearpage
\bibliographystyle{siam}
\bibliography{journal_list_long,bibliography,references}

>>>>

Figure \ref{fig:1dmasks} shows the simplest way to compute derivatives, but Roberts \cite{roberts1963machine} proposes
to compute them as shown in Figure \ref{fig:roberts}. In order to use the same convolution code for all algorithms, our implementation uses $3\times3$ matrices, adding a row and a column of zeros to the matrices in Figure \ref{fig:roberts}, what produces the same result.

@@

%-------------------------------------------------------------------------------
\clearpage
\bibliographystyle{siam}
\bibliography{bibliography}
